Enrich Object Detection : 2D-3D registration and continuous viewpoint estimation

### 2014-11-13 egpaper_final.tex
started intro pass
my change
Sam: Related works section done(ish)

\section{Related Works}
\subsection{Whitened Histograms of Orientations(WHO)}
% History of Whitening and how whitening works
Whitening a signal has been a popular method in various signal processing communities as a preprocessing stage \cite{Martinsson05, Belouchrani00}. Whitening a signal makes its autocorrelation narrower which results in more delta-function like spike for true positives and zero elsewhere. More formally, suppose that we have a $k$-dimensional random variable $X \in \mathcal{R}^k$ with $Cov(X)=\Sigma$. By whitening the signal, i.e., $\tilde{X}=\Sigma^{-\frac{1}{2}}(X - E[X])$, we remove 2nd order correlation between all features. So, all elements in the random variable $\tilde{X} = \left[\tilde{X}_1, \tilde{X}_2, \dots, \tilde{X}_k\right]$, become more independent. Since they become less dependent, convolving the whitened template with random signal will produce Gaussian noise on background signal but will produce delta-function like response on the match. Recently, \cite{Hariharan12} used whitening on HOG feature and named the whitened HOG as Whitened Histograms of Orientations (WHO).

@@

\subsection{More than a Bounding Box}

High performance object detectors generalize very well \cite{Felzenszwalb10, Girshick14}. They can handle intraclass variability, change in viewpoint and truncation. However due this generalization, they lose fine-grained information such as accurate pose and object sub-category.

To give more than just a rectangular bounding box, \cite{Pepik12, Xiang12, Fidler12, Xiang14, Hejrati14, Aubry14, Lim14} various methods exist to jointly detect and estimate the pose of an object. \cite{Xiang12, Hejrati14, Fidler12} used 3D representations that deform as viewpoint changes and \cite{Pepik12} uses geometric constraints to regularize 2D appearance models.

On fine-grained pose estimation, \cite{Zia13} uses a 3D deformable wireframe model with parts to accurately predict pose and align the model.

More recently, \cite{Aubry14, Lim14} showed a success in joint instance-level object detection and pose estimation. To estimate pose, they rely on synthetic renderings of CAD models and learn discriminative mid-level patches. \cite{Aubry14} finds discriminative mid-level patches and calibrates them on a small set of real images. Similarly, \cite{Lim14} finds and learns relative discriminativeness of the shared mid-level patches.

@@

We also used iterative Conjugate Gradient to make use of low rank property of the covariance \cite{Gharbi12}. After adding small regularization, the regularized covariance matrix has nice spectral property (Figure \ref{fig:spectral}. The singular values are clustered in small region and most of the energies are concentrated on the first (number) singular values. Thus, the conjugate gradients stage doesn't go upto the size of the matrix to converge.

\begin{figure}[t]
  \begin{center}
  \begin{tabular}{cc}
     \includegraphics[width=0.5\linewidth]{whotime} & 
     \includegraphics[width=0.5\linewidth]{speedup}\\
     (a) & (b) \\
 \end{tabular}
  \end{center}
  \caption{Comparison of HOG variants generation time and final }
  \label{fig:whotime}
\end{figure}
[Figure spectral analysis]

\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\linewidth]{residual}
  \caption{Residual of differentd method}
  \label{fig:whotime}
\end{figure}

@@

\subsection{Object Detection} 
Instance-level 2D-3D matching (registration) is a difficult problem since we need a CAD model that looks similar to the object in the image. Especially, due to extreme intraclass variability of chair \cite{Aubry14} used a very large collection of CAD models to cover variety of chairs. Similarly, \cite{Lim14}
\begin{table}[!htbp]
  \begin{center}
    \begin{tabular}{|c|c|c|c|}
    \hline
     AP/MPPE& Ours & ALM\cite{Xiang12} & VOCDPM\cite{Pepik12} \\
    \hline\hline
    car & \textbf{99.8}/91.7 &  98.4/93.4 & \textbf{99.8}/\textbf{97.5} \\ 
    bicycle & 93.0/87.4 & 93.0/91.4 & \textbf{98.8}/\textbf{97.5} \\
    \hline
    \end{tabular}
  \end{center}
  \caption{Average Precision(AP) and Average Viewpoint Accuracy(AVA) on 3DObject dataset. Though we do not need any training images and made templates in few minues, it performs on par with state of the art detectors yet it gives high quality 2D-3D matching.}
  \label{tab:3dobject}
\end{table}

>>>>

\section{Related Works}
\subsection{Whitened Histograms of Orientations (WHO)}
% History of Whitening and how whitening works
Recently, Hariharan \emph{et al.} introduced Whitened Histograms of Orientations (WHO) \cite{Hariharan12}, which used statistics from natural images to whiten HOG templates, making them more discriminative. Whitening is a common signal processing operation for decorrelating a set of random variables \cite{Martinsson05, Belouchrani00}. More formally, suppose that we have a $k$-dimensional random variable $X \in \mathcal{R}^k$ with $Cov(X)=\Sigma$. By whitening the signal, i.e., $\tilde{X}=\Sigma^{-\frac{1}{2}}(X - E[X])$, we remove 2nd order correlation between all features. So, all elements in the random variable $\tilde{X} = \left[\tilde{X}_1, \tilde{X}_2, \dots, \tilde{X}_k\right]$, become more independent. In the context of HOG features, a whitened HOG template will produce Gaussian noise when convolved with a random natural image, but a delta-like response on an image matching the template. WHO templates are more discriminative than standard HOG templates because they only capture the gradients that differentiate the object from random images.

\textcolor{red}{Should we mention how our method improves on this (ie by using CG)?}

@@

\subsection{Pose Detection}

Modern object detectors generalize very well, handling intraclass variability, occlusion, truncation and viewpoint changes \cite{Felzenszwalb10, Girshick14}. However, this generalization comes at the cost of fine-grained information, including accurate 3D pose and object sub-category recognition. Such methods typically produce bounding box detection hypotheses, will little further information.

Many methods have attempted to move object detection towards richer outputs, especially by jointly performing detection and pose estimation \cite{Pepik12, Xiang12, Fidler12, Xiang14, Hejrati14, Aubry14, Lim14}. To achieve this, \cite{Xiang12, Hejrati14, Fidler12} use 3D representations that deform as viewpoint changes and \cite{Pepik12} uses geometric constraints to regularize 2D appearance models.

The methods above perform discrete pose estimation, quantizing the viewing sphere into a number of poses and selecting the best one during inference. Fine-grained pose estimators, in contrast, can infer continuous (or arbitrarily fine-grained) poses. One such method from \cite{Zia13} aligns a 3D deformable part-based wireframe model with input images to accurately predict object poses.

More recently, \cite{Aubry14, Lim14} made progress in joint instance-level object detection and pose estimation. To estimate pose they use synthetic renderings of CAD models to learn discriminative mid-level patches. \cite{Aubry14} calibrates these patches on a small set of real images, while \cite{Lim14} present a method for learning the relative discriminativeness of the patches.

\textcolor{red}{Again, should we highlight the differences between these methods and ours?}

@@

We also used iterative Conjugate Gradient to make use of low rank property of the covariance \cite{Gharbi12}. After adding small regularization, the regularized covariance matrix has nice spectral property. figure below. The singular values are clustered in small region and most of the energies are concentrated on the first (number) singular values. Thus, the conjugate gradients stage doesn't go upto the size of the matrix to converge.

\begin{figure}[t]
  \begin{center}
  \begin{tabular}{cc}
     \includegraphics[width=0.5\linewidth]{whotime} & 
     \includegraphics[width=0.5\linewidth]{speedup}\\
     (a) & (b) \\
 \end{tabular}
  \end{center}
  \caption{Comparison of HOG variants generation time and final }
  \label{fig:whotime}
\end{figure}
[Figure spectral analysis]
[Figure residuals of each method]

@@

\subsection{Object Detection} 
Instance-level 2D-3D matching (registration) is a difficult problem since we need a CAD model that looks similar to the object in the image. Especially, due to extreme intraclass variability of chair \cite{Aubry14} used a very large collection of CAD models to cover variety of chairs. Similarly, \cite{Lim14} 
\begin{table}[!htbp]
  \begin{center}
    \begin{tabular}{|c|c|c|c|}
    \hline
     AP/AVA& Ours & ALM\cite{Xiang12} & VOCDPM\cite{Pepik12} \\
    \hline\hline
    car & \textbf{99.8}/91.7 &  98.4/93.4 & \textbf{99.8}/\textbf{97.5} \\ 
    bicycle & 93.0/87.4 & 93.0/91.4 & \textbf{98.8}/\textbf{97.5} \\
    \hline
    \end{tabular}
  \end{center}
  \caption{Average Precision(AP) and Average Viewpoint Accuracy(AVA) on 3DObject dataset. Though we do not need any training images and made templates in few minues, it performs on par with state of the art detectors yet it gives high quality 2D-3D matching.}
  \label{tab:3dobject}
\end{table}
